---
output:
  pdf_document: default
  html_document: default
---
Hunter Dawley

# REQUIRED PACKAGES/LOAD META DATA
```{r}
library(sf) # classes and functions for vector data
library(raster) # classes and functions for raster data
library(spData) # load geographic data
library(stars)
library(gapminder)
library(mapdata)
library(devtools)
library(tidyverse) 
library(ggmap)
library(dplyr)


parks <- read_csv("../../2_data/raw_data/parks.csv")
View(parks)

species <- read_csv("../../2_data/raw_data/species.csv")
view(species)

#Looking into these national parks CALIFORNIA: Redwoods (REDW), Yosemite (YOSE), Pinnacles (PINN), Lassen Volcanic (LAVO), Channel Island (CHIS), Death Valley (DEVA), Joshua Tree (JOTR)

#COLORADO: Black Canyon of the Gunnison (BLCA), Great Sand Dunes (GRSA), Mesa Verde (MEVE), Rocky Mountain (ROMO)

#FLORIDA: Biscayne (BISC), Dry Tortugas (DRTO), Everglades (EVER)
```


#select observations for only the national parks we are looking into for the parks and species dataset
```{r}
#CALIFORNIA: Redwoods (REDW), Yosemite (YOSE), Pinnacles (PINN), Lassen Volcanic (LAVO), Channel Island (CHIS), Death Valley (DEVA), Joshua Tree (JOTR)

#COLORADO: Black Canyon of the Gunnison (BLCA), Great Sand Dunes (GRSA), Mesa Verde (MEVE), Rocky Mountain (ROMO)

#FLORIDA: Biscayne (BISC), Dry Tortugas (DRTO), Everglades (EVER)

parks_new <- parks %>%
  filter(`Park Name` =="Pinnacles National Park"|`Park Name` == "Redwood National Park" | `Park Name` == "Yosemite National Park" | `Park Name` == "Lassen Volcanic National Park" | `Park Name` == "Channel Islands National Park" | `Park Name` =="Death Valley National Park" | `Park Name` =="Joshua Tree National Park"| `Park Name` =="Black Canyon of the Gunnison National Park"| `Park Name` =="Great Sand Dunes National Park and Preserve"| `Park Name` == "Mesa Verde National Park"|`Park Name` == "Rocky Mountain National Park"|`Park Name` == "Biscayne National Park"|`Park Name` =="Dry Tortugas National Park"| `Park Name` == "Everglades National Park")

species_new <- species %>%
  filter(`Park Name` =="Pinnacles National Park"|`Park Name` == "Redwood National Park" | `Park Name` == "Yosemite National Park" | `Park Name` == "Lassen Volcanic National Park" | `Park Name` == "Channel Islands National Park"| `Park Name` =="Death Valley National Park" | `Park Name` =="Joshua Tree National Park"| `Park Name` =="Black Canyon of the Gunnison National Park"| `Park Name` =="Great Sand Dunes National Park and Preserve"| `Park Name` == "Mesa Verde National Park"|`Park Name` == "Rocky Mountain National Park"|`Park Name` == "Biscayne National Park"|`Park Name` =="Dry Tortugas National Park"| `Park Name` == "Everglades National Park")
```

#look only at mammals and birds for the species dataset
```{r}
species_2 <- species_new %>%
  filter(`Category` == "Mammal" | `Category` == "Bird")

str(species_2)
str(parks_new)

species_2 %>%
  group_by(Category) %>%
  summarise(n=n()) 

#get counts of birds and mammals per park
species_final <- species_2 %>%
  group_by(Category, `Park Name`) %>%
  summarise(n = n()) %>%
  rename(park = `Park Name`)

species_final
```

#rename columns to be able to join
```{r}
species_final$park <- recode(species_final$park, 
                             `Biscayne National Park` = "Biscayne",
                             `Black Canyon of the Gunnison National Park` = "Black Canyon",
                             `Channel Islands National Park` = "Channel Islands",
                             `Death Valley National Park` = "Death Valley",
                             `Dry Tortugas National Park` = "Dry Tortugas",
                             `Everglades National Park` = "Everglades",
                             `Great Sand Dunes National Park and Preserve` = "Great Sand Dunes",
                             `Joshua Tree National Park` = "Joshua Tree",
                             `Lassen Volcanic National Park` = "Lassen Volcanic",
                             `Mesa Verde National Park` = "Mesa Verde",
                             `Pinnacles National Park` = "Pinnacles",
                             `Redwood National Park` = "Redwoods",
                             `Rocky Mountain National Park` = "Rocky Mountain",
                             `Yosemite National Park`= "Yosemite"
                              )


parks_new$`Park Name` <- recode(parks_new$`Park Name`, 
                             `Biscayne National Park` = "Biscayne",
                             `Black Canyon of the Gunnison National Park` = "Black Canyon",
                             `Channel Islands National Park` = "Channel Islands",
                             `Death Valley National Park` = "Death Valley",
                             `Dry Tortugas National Park` = "Dry Tortugas",
                             `Everglades National Park` = "Everglades",
                             `Great Sand Dunes National Park and Preserve` = "Great Sand Dunes",
                             `Joshua Tree National Park` = "Joshua Tree",
                             `Lassen Volcanic National Park` = "Lassen Volcanic",
                             `Mesa Verde National Park` = "Mesa Verde",
                             `Pinnacles National Park` = "Pinnacles",
                             `Redwood National Park` = "Redwoods",
                             `Rocky Mountain National Park` = "Rocky Mountain",
                             `Yosemite National Park`= "Yosemite"
                              )
```                            

#visualizations of biodiviersity in the parks
```{r}
#CALIFORNIA BIRDS AND MAMMALS AND GRAPHS
CA_Biodiversity <- species_final %>%
  filter(`park` == "Yosemite" | `park` == "Redwoods" | `park` == "Joshua Tree" | `park` == "Pinnacles" | `park` == "Lassen Volcanic" | `park` == "Channel Islands" | `park` == "Death Valley")
CA_biodiversity_visual <- ggplot(CA_Biodiversity, aes(x = Category, y = n, fill = Category)) + 
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = n), vjust = 0) +
  facet_wrap(~park) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 600)) + 
  theme_minimal()
  ggtitle("California Bird/Mammal Biodiversity") 

CA_biodiversity_visual
ggsave("CA_biodiversity_visual.png", path = "../clean_data")

#FLORIA PARKS
FL_Biodiversity <- species_final %>%
  filter(`park` == "Dry Tortugas" | `park` =="Everglades" | `park` == "Biscayne")
FL_biodiversity_visual <- ggplot(FL_Biodiversity, aes(x = Category, y = n, fill = Category)) + 
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = n), vjust = 0) +
  facet_wrap(~park) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 600))+
  ggtitle("Florida Bird/Mammal Biodiversity") + 
  theme_minimal()
ggsave("FL_biodiversity_visual.png", path = "../clean_data")

#COLORADO PARKS
CO_Biodiversity <- species_final %>%
  filter(`park` == "Black Canyon" | `park` == "Great Sand Dunes"| `park` == "Mesa Verde" | `park` == "Rocky Mountain") 
CO_biodiversity_visual <- ggplot(CO_Biodiversity, aes(x = Category, y = n, fill = Category)) + 
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = n), vjust = 0) +
  facet_wrap(~park) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 600))+
  ggtitle("Colorado Bird/Mammal Biodiversity") +
  theme_minimal()
ggsave("CO_biodiversity_visual.png", path = "../clean_data")

#ALL NATIONAL PARK GRAPHS TOGETHER
all_park_biodiversity <- ggplot(species_final, aes(x = Category, y = n, fill = Category)) + 
  geom_bar(stat = "identity", width = 1) +
  geom_text(aes(label = n), vjust = 0) +
  facet_wrap(~park) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 600)) +
  ggtitle("All Park Bird/Mammal Biodiversity") 
ggsave("all_park_biodiversity.png", path = "../clean_data")
```
#find the average of birds and mammals in each park for each state
```{r}
CA_Biodiversity_total <- CA_Biodiversity %>%
  group_by(Category) %>%
  summarise(total_biodiversity = sum(n)) %>%
  mutate(avg_biodiversity = round(total_biodiversity/7, 2))
#on average, CA parks have 103.57 mammals and 339.29 birds per national park

CO_Biodiversity_total <- CO_Biodiversity %>%
  group_by(Category) %>% 
  summarise(total_biodiversity = sum(n)) %>%
  mutate(avg_biodiversity = round(total_biodiversity/4, 2))
#on average, CO parks have 73.75 mammals and 232.25 birds per national park

FL_Biodiversity_total <- FL_Biodiversity %>%
  group_by(Category) %>%
  summarise(total_biodiversity = sum(n)) %>% 
  mutate(avg_biodiversity = round(total_biodiversity/3, 2))
#on average, fl parks have 25.33 mammals and 295.33 birds per national park

```


# join the parks_new set with the species_final
```{r}
join_species_park <- full_join(parks_new, species_final, by = c('Park Name' ='park'))
view(join_species_park)
```

# join join_species_park with the weather data (called weather biodiversity)
```{r}
#we only have one city weather data for Colorado (Denver) and all of the FL national parks are closest to Miami so we will use the same data.

#Redwoods, Yosemite and Pinnacles, Lassen Volcanic are San Fran. Channel Islands and Joshua Tree are closest to Los Angeles 

join_species_park$city <- c("Miami", "Miami" , "Denver", "Denver", "Los Angeles", "Los Angeles", "Los Angeles", "Los Angeles", "Miami", "Miami" , "Miami", "Miami", "Denver", "Denver", "Los Angeles", "Los Angeles", "San Francisco", "San Francisco", "Denver", "Denver", "San Francisco", "San Francisco", "San Francisco", "San Francisco", "Denver", "Denver", "San Francisco", "San Francisco" )

join_weather_2017 <- read_csv("join_weather_2017.csv", show_col_types = FALSE)

weather_biodiversity <- full_join(join_weather_2017, join_species_park, by=c("State_ini"="State"))
```


#need to find average temperature for each park for each month
```{r}
park_temps <- weather_biodiversity %>%
  drop_na() %>%
  group_by(`Park Name`, `month`) %>%
  summarise(Monthly_temp = mean(temperature, na.rm = TRUE))

#NA values for Death Valley temps-drop na
#need to make a column of closest city to each national park. We only have select city weather data.
```


#visualization of weather in all the parks
```{r}
ggplot(park_temps, aes(x = month, y = Monthly_temp, fill = `Park Name`)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~`Park Name`) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 100)) 

#FLORIA PARKS
FL_Biodiversity_weather <- park_temps %>%
filter(`Park Name` == "Dry Tortugas" | `Park Name` =="Everglades" | `Park Name` == "Biscayne")

ggplot(FL_Biodiversity_weather, aes(x = month, y = Monthly_temp, fill = `Park Name`)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~`Park Name`) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 100)) 

#COLORADO PARKS
CO_Biodiversity_weather <- park_temps %>%
filter(`Park Name` == "Black Canyon" | `Park Name` =="Great Sand Dunes" | `Park Name` == "Mesa Verde" | `Park Name` == "Rocky Mountain")

ggplot(CO_Biodiversity_weather, aes(x = month, y = Monthly_temp, fill = `Park Name`)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~`Park Name`) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 100))  

#CALIFORNIA PARKS
CA_Biodiversity_weather <- park_temps %>%
  filter(`Park Name` == "Yosemite" | `Park Name` == "Redwoods" | `Park Name` == "Joshua Tree" | `Park Name` == "Pinnacles" | `Park Name` == "Lassen Volcanic" | `Park Name` == "Channel Islands"| `Park Name` == "Death Valley")

ggplot(CA_Biodiversity_weather, aes(x = month, y = Monthly_temp, fill = `Park Name`)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~`Park Name`) + 
  scale_fill_brewer(palette = "Paired") + 
  scale_y_continuous(name="Count", limits=c(0, 100)) 

#only monthly temps until october for San Fran and Miami.
```


```{r}


write.csv(weather_biodiversity, file = "final_weather_biodiversity", row.names = FALSE)
#use this!!

#write.csv(join_species_park, file = "join_species_park.csv", row.names = FALSE)
#write.csv(park_temps, file = "park_temps.csv", row.names = FALSE)
```


#USE A MAP TO VISUALIZE BIODIVIERSITY
```{r}
bird_biodiversity <- weather_biodiversity %>% 
  filter(Category == "Bird") 

bird_biodiversity_plot <- bird_biodiversity %>%
  ggplot(aes(x = Longitude, y = Latitude)) +
    borders("state") +
    xlim(-125, -65)+ ylim(25,50)+
    scale_colour_gradient(low = "sky blue", high = "black")+
    geom_point(aes(color = n), size = 3) +
    coord_quickmap()+
    labs(title = "2017 Bird Biodiversity Among Parks", 
         colour="Birds", x ="Longitude")+
  theme_minimal()

bird_biodiversity_plot
ggsave("bird_biodiversity.png", path = "../clean_data")
```

```{r}
mammal_biodiversity <- weather_biodiversity %>% 
  filter(Category == "Mammal") 

mammal_biodiversity_plot <- mammal_biodiversity %>%
  ggplot(aes(x = Longitude, y = Latitude)) +
    borders("state") +
    xlim(-125, -65)+ ylim(25,50)+
    scale_colour_gradient(low = "sky blue", high = "black")+
    geom_point(aes(color = n), size = 3) +
    coord_quickmap()+
    labs(title = "2017 Mammal Biodiversity Among Parks", 
         colour="Mammals", x ="Longitude")+
  theme_minimal()
mammal_biodiversity_plot
ggsave("mammal_biodiversity.png", path = "../clean_data")
```



