---
title: "data-cleaning-JiNoh"
author: "Ji Noh"
date: "2022-11-03"
output: github_document
---

# Data Cleaning for Weather datasets



```{r}
library(dplyr)
library(tidyverse)
```


## city_attributes dataset


Here, I will first load the city attributes dataset. This data contains "City". "Country", "Latitude", and "Longitude" information.

```{r}
#load dataset

library(readr)
city_attributes <- read_csv("../../2_data/raw_data/city_attributes.csv")

```


Here, I will make a new data called *city_attributes_clean* to have observations that we only need for your analysis.

I am going to filter cities of *"San Francisco", "Los Angeles", "San Diego", "Denver", "Jacksonville", and "Miami"*.


```{r}
#filtering cities we need

city_attributes_clean <- city_attributes %>%
  filter(City == "San Francisco" | City == "Los Angeles" | City == "San Diego" | City == "Denver" | City == "Jacksonville" | City == "Miami")

city_attributes_clean
```

Now, we will save this dataset as a separate csv file called *"city_lat_and_long"*.
```{r}
#write.csv(city_attributes_clean, file="city_lat_and_long.csv", row.names = FALSE)
```


## Humidity data

```{r}
#Load data
library(readr)
humidity <- read_csv("../../2_data/raw_data/humidity.csv")

```


Here, unlike *city_attributes* dataset, each city is in the column. I will do pivot_longer() to go that into observation rather than being variable names.

```{r}
#pivot_longer() to make humidity and city as observations.
head(humidity)
humidity_clean <- humidity %>%
  select("datetime", "San Francisco", "Los Angeles", "San Diego", "Denver", "Jacksonville", "Miami") %>%
  pivot_longer("San Francisco":"Miami", names_to = "City", values_to = "Humidity")  %>%
  filter(grepl('2017',datetime)) 


humidity_clean
```


Looking at the *datetime* variable, it is formed as year-month-day-time. We might want to split these to make an analysis based on time or day or month. Here I will use format() function to split the *datatime* variable.

```{r}

#spliting date and time
humidity_clean$time <- format(as.POSIXct(humidity_clean$datetime), format = "%H:%M:%S")
humidity_clean$date <- as.Date(humidity_clean$datetime) 

#spliting year month day
humidity_clean <- humidity_clean %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))

humidity_clean

```
Now we have variables of *"year", "Humidity", "City", "day", "month", "time", "datetime", and "date"*.
I am still leaving the original *"datetime"* variable just in case for later use.

Next, I would like to arrange the dataset by *City"*, to make it more clear.

```{r}
humidity_clean <- humidity_clean %>%
  arrange(City)
```

I am saving this *humidity_clean* dataset to csv file format.

```{r}
#write.csv(humidity_clean, file="humidity_by_hour_clean.csv", row.names = FALSE)
```


## pressure data

```{r}
#load data
library(readr)
pressure <- read_csv("../../2_data/raw_data/pressure.csv")

```
 
I will first filter out the cities we want, and save it to a new dataset called pressure_clean.

```{r}
pressure_clean <- pressure %>%
  select("datetime", "San Francisco", "Los Angeles", "San Diego", "Denver", "Jacksonville", "Miami")

pressure_clean
```

Here, similar with humidity dataset, the cities are in the variables. I will use the same method I used for the humidity to make the cities and pressure values to go into the observation, using pivot_longer() function.

I will also make new variables called *"Pressure"* and *"City"* to store all the observations there.

```{r}
#pivot_longer() to make pressure and cities to go into observation.
pressure_clean <- pressure %>%
  select("datetime", "San Francisco", "Los Angeles", "San Diego", "Denver", "Jacksonville", "Miami") %>%
  pivot_longer("San Francisco":"Miami", names_to = "City", values_to = "Pressure")  %>%
  filter(grepl('2017',datetime)) 



#spliting date and time
pressure_clean$time <- format(as.POSIXct(pressure_clean$datetime), format = "%H:%M:%S")
pressure_clean$date <- as.Date(pressure_clean$datetime) 

#spliting year month day
pressure_clean <- pressure_clean %>%
  dplyr::mutate(year = lubridate::year(date), 
                month = lubridate::month(date), 
                day = lubridate::day(date))

pressure_clean

```

After discussing with teammates, we decided not to use pressure information for our analysis. Our team is going to look at **temperature**, **humidity**, and **wind speed** to see if there is any relationships with weather, visitations, and biodiversity. Therefore, for the pressure data, I will not going to save this to separate csv file.

