---
title: "viz for preprocessing"
author: "Ji Noh"
date: "2022-12-08"
output: html_document
---

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

## Loading the data from master data
```{r}
library(readr)
df <- read_csv("https://www.dropbox.com/s/afzhasp96hvli1t/master_12_17.csv?dl=1")

df
```

## 1. Temperature data

### Simple findings
```{r}
temperature_avg <- df %>%
  select(City,Year,Month,day,time,temperature, avg_temperature, ParkName) %>%
  group_by(Month, City) %>%
  summarise(avg_temp_year = mean(avg_temperature, na.rm = TRUE))



```


```{r}
oc.color <- c("#2196F3", "#9FA8DA", "#4DD0E1","#81C784")

temp_plot <-ggplot(temperature_avg, aes(x=Month, y= avg_temp_year, color = City)) +
  geom_line()+
  labs(title = "2012-2017 Average Temperature by Month For Cities",
       x = "Month",
       y = "Average Humidity") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  scale_fill_manual(values = oc.color[1:4]) +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) +
  theme_minimal() 

ggsave(filename = "temp_general.png",
       plot = temp_plot,
       width = 10,
       height = 7)
```


### Specific findings


First, we would like to see all the temperature in our desired cities from 2012-2017.
I will then group them with month to see the temperature patterns in each cities by month.


```{r}
#selecting the temperature variable
temperature <- df %>%
  select(City,Year,Month,day,time,temperature, ParkName)
```

```{r}
#denver
temp_data_denver <- temperature %>%
  filter(City == "Denver") %>%
  select(City:temperature)

#temp_data_denver

temp_denver_plot <- ggplot(temp_data_denver, aes(x=Month, y = temperature, group=Month)) +
  geom_boxplot(fill="lightblue2") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 Denver Temperature by Month",
    x = "Month",
    y = "Temperature",
    fill = "Temperature"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#temp_denver_plot
```

```{r}
#Miami
temp_data_miami <- temperature %>%
  filter(City == "Miami") %>%
  select(City:temperature)

#temp_data_miami

temp_miami_plot <- ggplot(temp_data_miami, aes(x=Month, y = temperature, group=Month)) +
  geom_boxplot(fill="lightcyan2") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 Miami Temperature by Month",
    x = "Month",
    y = "Temperature",
    fill = "Temperature"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#temp_miami_plot
```

```{r}
#Los Angeles
temp_data_la<- temperature %>%
  filter(City == "Los Angeles") %>%
  select(City:temperature)

#temp_data_la

temp_la_plot <- ggplot(temp_data_la, aes(x=Month, y = temperature, group=Month)) +
  geom_boxplot(fill="lightsteelblue") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 Los Angeles Temperature by Month",
    x = "Month",
    y = "Temperature",
    fill = "Temperature"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#temp_la_plot
```

```{r}
#San Francisco
temp_data_sf<- temperature %>%
  filter(City == "San Francisco") %>%
  select(City:temperature)

#temp_data_sf

temp_sf_plot <- ggplot(temp_data_sf, aes(x=Month, y = temperature, group=Month)) +
  geom_boxplot(fill="paleturquoise3") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 San Francisco Temperature by Month",
    x = "Month",
    y = "Temperature",
    fill = "Temperature"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#temp_sf_plot
```


```{r}
#then, I will arrange above four plots
library(gridExtra)
grid.arrange(temp_denver_plot,temp_miami_plot,temp_la_plot,temp_sf_plot)

#ggsave(filename = "Temperature_compare_by_city_3.png",
      # grid.arrange(temp_denver_plot,temp_miami_plot,temp_la_plot,temp_sf_plot),
      # width = 30,
       #height = 20)
```






## 2. Humidity data

### Simple findings
```{r}
hum_avg <- df %>%
  select(City,Year,Month,day,time,humidity, avg_humidity, ParkName) %>%
  group_by(Month, City) %>%
  summarise(avg_hum_year = mean(avg_humidity, na.rm=TRUE))

```


```{r}
oc.color <- c("#2196F3", "#9FA8DA", "#4DD0E1","#81C784")

hum_plot <-ggplot(hum_avg, aes(x=Month, y= avg_hum_year, color = City)) +
  geom_line()+
  labs(title = "2012-2017 Average Humidity by Month For Cities",
       x = "Month",
       y = "Average Humidity") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  scale_fill_manual(values = oc.color[1:4]) +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) +
  theme_minimal() 

ggsave(filename = "hum_general.png",
       plot = hum_plot,
       width = 10,
       height = 7)
```


### Specific findings

Here, we will do the same process as above for the humidity observations.

```{r}
hum <- df %>%
  select(City,Year,Month,day,time,humidity, ParkName) 

hum
```


```{r}
#denver
hum_data_denver <- hum %>%
  filter(City == "Denver") %>%
  select(City:humidity)

#hum_data_denver

hum_denver_plot <- ggplot(hum_data_denver, aes(x=Month, y = humidity, group=Month)) +
  geom_boxplot(fill="lightblue2") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 Denver Humidity by Month",
    x = "Month",
    y = "Humidity",
    fill = "Humidity"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 
#hum_denver_plot

```

```{r}
#Miami
hum_data_miami <-  hum %>%
  filter(City == "Miami") %>%
  select(City:humidity)

#hum_data_miami

hum_miami_plot <- ggplot(hum_data_miami, aes(x=Month, y = humidity, group=Month)) +
  geom_boxplot(fill="lightcyan2") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 Miami Humidity by Month",
    x = "Month",
    y = "Humidity",
    fill = "Humidity"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#hum_miami_plot
```

```{r}

#Los Angeles
hum_data_la<- hum %>%
  filter(City == "Los Angeles") %>%
  select(City:humidity)

#hum_data_la

hum_la_plot <- ggplot(hum_data_la, aes(x=Month, y = humidity, group=Month)) +
  geom_boxplot(fill="lightsteelblue") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 Los Angeles Humidity by Month",
    x = "Month",
    y = "Humidity",
    fill = "Humidity"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#hum_la_plot
```


```{r}
#San Francisco
hum_data_sf<- hum %>%
  filter(City == "San Francisco") %>%
  select(City:humidity)

#hum_data_sf

hum_sf_plot <- ggplot(hum_data_sf, aes(x=Month, y = humidity, group=Month)) +
  geom_boxplot(fill="paleturquoise3") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,100,20),  limits=c(0, 100)) +
  labs(
    title = "2012 - 2017 San Francisco Humidity by Month",
    x = "Month",
    y = "Humidity",
    fill = "Humidity"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#hum_sf_plot
```


```{r}
#alligning all the plots
library(gridExtra)
grid.arrange(hum_denver_plot,hum_miami_plot,hum_la_plot,hum_sf_plot)

#ggsave(filename = "humidity_compare_by_city1.png",
      # plot = grid.arrange(hum_denver_plot,hum_miami_plot,hum_la_plot,hum_sf_plot),
      # width = 30,
      # height = 20)
```


## 3. Wind Speed data

### Simple findings
```{r}
wind_avg <- df %>%
  select(City,Year,Month,day,time,wind_speed, avg_wind_speed, ParkName) %>%
  group_by(Month, City) %>%
  summarise(avg_wind_year = mean(avg_wind_speed, na.rm = TRUE))

```


```{r}
oc.color <- c("#2196F3", "#9FA8DA", "#4DD0E1","#81C784")

wind_plot <- ggplot(wind_avg, aes(x=Month, y= avg_wind_year, color = City)) +
  geom_line() +
  labs(title = "2012-2017 Average Wind Speed by Month For Cities",
       x = "Month",
       y = "Average Wind Speed") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,5,1),  limits=c(0, 5)) +
  scale_fill_manual(values = oc.color[1:4]) +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) +
  theme_minimal() 

ggsave(filename = "wind_general.png",
       plot = wind_plot,
       width = 10,
       height = 7)
```



Sampe process with wind speed observations
```{r}
wind <- df %>%
  select(City,Year,Month,day,time,wind_speed, ParkName) 

wind
```

```{r}
#denver
wind_data_denver <- wind %>%
  filter(City == "Denver") %>%
  select(City:wind_speed)

#wind_data_denver

wind_denver_plot <- ggplot(wind_data_denver, aes(x=Month, y = wind_speed, group=Month)) +
  geom_boxplot(fill="lightblue2") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,30,5),  limits=c(0, 30)) +
  labs(
    title = "2012 - 2017 Denver Wind Speed by Month",
    x = "Month",
    y = "Wind Speed",
    fill = "Wind Speed"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 
#wind_denver_plot

```

```{r}
#Miami
wind_data_miami <-  wind %>%
  filter(City == "Miami") %>%
  select(City:wind_speed)

#wind_data_miami

wind_miami_plot <- ggplot(wind_data_miami, aes(x=Month, y = wind_speed, group=Month)) +
  geom_boxplot(fill="lightcyan2") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,30,5),  limits=c(0, 30)) +
  labs(
    title = "2012 - 2017 Miami Wind Speed by Month",
    x = "Month",
    y = "Wind Speed",
    fill = "Wind Speed"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#wind_miami_plot

```


```{r}
#Los Angeles
wind_data_la<- wind %>%
  filter(City == "Los Angeles") %>%
  select(City:wind_speed)

#wind_data_la

wind_la_plot <- ggplot(wind_data_la, aes(x=Month, y = wind_speed, group=Month)) +
  geom_boxplot(fill="lightsteelblue") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,30,5),  limits=c(0, 30)) +
  labs(
    title = "2012 - 2017 Los Angeles Wind Speed by Month",
    x = "Month",
    y = "Wind Speed",
    fill = "Wind SPeed"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#wind_la_plot
```

```{r}
#San Francisco
wind_data_sf<- wind %>%
  filter(City == "San Francisco") %>%
  select(City:wind_speed)

#wind_data_sf

wind_sf_plot <- ggplot(wind_data_sf, aes(x=Month, y = wind_speed, group=Month)) +
  geom_boxplot(fill="paleturquoise3") +
  scale_x_continuous(breaks = seq(0,12,1)) +
  scale_y_continuous(breaks = seq(0,30,5),  limits=c(0, 30)) +
  labs(
    title = "2012 - 2017 San Francisco Wind Speed by Month",
    x = "Month",
    y = "Wind Speed",
    fill = "Wind Speed"
  ) +
  theme_minimal() +
  theme(plot.title=element_text(size=40),
        axis.title=element_text(size=30),
        axis.text =element_text(size=30)) 

#wind_sf_plot
```

```{r}
library(gridExtra)
grid.arrange(wind_denver_plot,wind_miami_plot,wind_la_plot,wind_sf_plot)

ggsave(filename = "windspeed_compare_by_city.png",
       grid.arrange(wind_denver_plot,wind_miami_plot,wind_la_plot,wind_sf_plot),
       width = 30,
       height = 20)
```