---
title: "data visualization for map and time series"
output: github_document
---
```{r, message=FALSE}
library(gapminder)
library(mapdata)
library(tidyverse) 
library(ggmap)
library(readr)
library(Ecfun)
library(ggrepel)
# TS
library(grid) # for text in TS
library(lubridate)
library(zoo)
library(ggpubr)

# read in master dataset
weather <- read.csv("https://www.dropbox.com/s/afzhasp96hvli1t/master_12_17.csv?dl=1")
```

### Time Series
join the State name and create monthly weather data
Colorado's annual temperature fluctuates the most, followed by Florida's and California's, which fluctuate the least, according to the graph.
```{r, message=FALSE}
# convert date variables into Date format 
weather$YMdate <- as.Date(weather$YMdate)
weather$date <- as.Date(weather$date)
monthly <- weather %>% 
  select("ParkName", "State", "City","RecreationVisits", "avg_temperature","YMdate","avg_humidity","avg_wind_speed") %>% 
  unique()
  
# la:3; SF :4; Miami:3; Denver:4

by_cities <- monthly %>% 
  group_by(State, City,avg_temperature,avg_humidity,avg_wind_speed,  YMdate) %>% 
  summarise(avg_visits_city = mean(RecreationVisits, rm.na = TRUE))

```
#### 1. Monthly
```{r, message=FALSE, warning=FALSE}
# prepare for the texts for two kind of lines
## temperature
grob1 <- grobTree(textGrob("temperature", x=0,  y=0.94, hjust=0,
  gp=gpar(col="black", fontsize=11, fontface="italic", alpha = 0.5)))
grob2<- grobTree(textGrob("visitaion", x=0,  y=0.09, hjust=0,
  gp=gpar(col="black", fontsize=11, fontface="italic", alpha = 0.5)))

by_cities %>% 
ggplot(aes(x = YMdate, group = City, color = State)) +
  geom_line(aes(y = avg_visits_city/5000), size = 1)+
  geom_point(aes(y =avg_temperature),alpha = 0.8,size = 0.4) +
  geom_line(aes(y =avg_temperature), alpha = 0.5) +
  scale_y_continuous(name = "temperature(F°)",sec.axis = sec_axis(~.*5 , name="visitation (by 1K)"))+
  labs(title = "Monthly Temperature and Visitation Change in 3 States (2012 - 2017)",
       x = "time") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  facet_grid(rows = vars(City))+  
  theme_light() +
  theme(panel.grid.major = element_blank())+
  annotation_custom(grob1)+annotation_custom(grob2)
# ggsave("visit_temp.png")


## humidity
grob1 <- grobTree(textGrob("humidity", x=0,  y=0.93, hjust=0,
  gp=gpar(col="black", fontsize=11, fontface="italic", alpha = 0.5)))
grob2<- grobTree(textGrob("visitaion", x=0,  y=0.09, hjust=0,
  gp=gpar(col="black", fontsize=11, fontface="italic", alpha = 0.5)))

by_cities %>% 
ggplot(aes(x = YMdate, group = City, color = State)) +
  geom_point(aes(y =avg_humidity),alpha = 0.8,size = 0.4) +
  geom_line(aes(y =avg_humidity), alpha = 0.5) +
  geom_line(aes(y = avg_visits_city/5000), size = 1)+
  theme_light() +
  scale_y_continuous(name = "humidity(%)",sec.axis = sec_axis(~.*5 , name="visitation (by 1K)"))+
  labs(title = "Monthly Humidity and Visitation Change in 3 States (2012 - 2017)",
       x = "time") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  facet_grid(rows = vars(City))+ 
  theme(panel.grid.major = element_blank())+
  annotation_custom(grob1)+annotation_custom(grob2)
# ggsave("vist_humidity.png")


## wind speed
grob1 <- grobTree(textGrob("wind", x=0,  y=0.93, hjust=0,
  gp=gpar(col="black", fontsize=11, fontface="italic", alpha = 0.5)))
grob2<- grobTree(textGrob("visitaion", x=0,  y=0.09, hjust=0,
  gp=gpar(col="black", fontsize=11, fontface="italic", alpha = 0.5)))

by_cities %>% 
ggplot(aes(x = YMdate, group = City)) +
  geom_point(aes(y =avg_wind_speed, color = State),alpha = 0.8,size = 0.4) +
  geom_line(aes(y =avg_wind_speed, color = State), alpha = 0.5) +
  geom_line(aes(y = avg_visits_city/100000, color = State), size = 1)+
  theme_light() +
  scale_y_continuous(name = "windspeed",sec.axis = sec_axis(~.*10 , name="visitation (by 10K)"))+
  labs(title = "Monthly Wind Speed and Visitation Change in 3 States (2012 - 2017)",
       x = "time") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+   
  facet_grid(rows = vars(City))+  
  theme(panel.grid.major = element_blank())+
  annotation_custom(grob1)+annotation_custom(grob2)

# ggsave("vist_windspeed.png")

```

#### 2.Quarterly
Now let's see the quarterly temperature comparison between cities and the three states.
As we guessed, the average temperature of FL is the highest, followed by CA, and CO is the lowest.
```{r, message=FALSE, warning=FALSE}
weather_qtr <- weather %>% 
  mutate(quarter = quarter(date), year = year(date)) %>% 
  group_by(year, quarter,State, City) %>% 
  summarise(quarterly_temperature = round(mean(avg_temperature, na.rm = TRUE),2),
            quarterly_wind_speed = round(mean(avg_wind_speed, na.rm = TRUE),2),
            quarterly_humidity = round(mean(avg_humidity, na.rm = TRUE),2)) %>% 
  unite("tmp", c("year", "quarter"), sep = "-") %>% 
  mutate(date = as.Date(as.yearqtr(tmp)))

## temperature
p1 = ggplot(weather_qtr,aes(x =date, y = quarterly_temperature, group = City, color = State)) +
  geom_point(aes(shape=City), alpha = 0.8, size = 2, show.legend = FALSE) +
  geom_line(alpha = 0.5, show.legend = FALSE) +
  labs(title = "Quarterly Temperature Change in 4 cities",
       x = "year by quarter",
       y = "quarterly temperature") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  theme_minimal() 

## humidity
p2 = ggplot(weather_qtr,aes(x =date, y = quarterly_humidity, group = City, color = State)) +
  geom_point(aes(shape=City), alpha = 0.8, size = 2) +
  geom_line(alpha = 0.5) +
  labs(title = "Quarterly Humidity Change in 4 cities",
       x = "year by quarter",
       y = "quarterly humidity") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  theme_minimal() 

## windspeed
p3 = ggplot(weather_qtr,aes(x =date, y = quarterly_wind_speed, group = City, color = State)) +
  geom_point(aes(shape=City), alpha = 0.8, size = 2) +
  geom_line(alpha = 0.5) +
  labs(title = "Quarterly Wind Speed Change in 4 cities",
       x = "year by quarter",
       y = "quarterly wind speed") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")+
  theme_minimal() 
# ggsave("quarterly_windspeed.png", width = 6, height = 4)

re = ggarrange(p1, p2,p3, nrow = 2, ncol = 2)
re 
# ggsave("quarterly_TS.png",re, dpi = "retina",width = 12,limitsize = FALSE)
```

### Maps
there is no benefits containing all the data we'll use below in our master dataset, because our master dataset is large already and we will not use such data to do furtuer analysis.
```{r, message=FALSE, warning=FALSE}
# load in raw data because our mater data contains only 14 parks we are interested in
loc_park <- read.csv("parks.csv")
all_park <- read.csv("National Parks Visitation Data.csv")

all_park$RecreationVisits <- asNumericChar(all_park$RecreationVisits)

loc_park[loc_park$Park.Name == "Death Valley National Park","State"] = "CA"

park_vis_17 <- all_park %>% filter(Year == 2017) %>% 
  group_by(Year,ParkName, UnitCode, State) %>% 
  summarise(annual_vis = round(mean(RecreationVisits, rm.na = TRUE), 2)) %>% 
  left_join(loc_park, by = c("UnitCode" = "Park.Code", "State" = "State")) %>% 
  na.omit()
```


####park visitaion map in 2017
```{r, message=FALSE, warning=FALSE}
# regular maps
# comment filter to draw all parks in the states
 
park_vis_17 %>% 
  # filter(State %in% c("CA","CO","FL")) %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
    borders("state") +
    xlim(-125, -65)+ ylim(25,50)+
    scale_colour_gradient(low = "sky blue", high = "black")+
    geom_point(aes(color = annual_vis), size = 3) +
    coord_quickmap()+
    labs(title = "2017 Annual Park Visition", colour="visition", x ="Longitute")+
    theme_minimal()

# ggsave("all_park_map.png")
# ggsave("4_park_map.png")
```

#### google maps
please use your google map api key for this part, `google_api_key`
```{r, data preparation, message=FALSE, warning=FALSE}
google_api_key = "xxxx"
register_google(key = google_api_key) 

bio <- read.csv("https://www.dropbox.com/s/f9yfrim8h9g88xx/final_weather_biodiversity?dl=1")
new_bio_bird <- bio %>% 
  select("Park.Code", "Park.Name", "State","Latitude", "Longitude", "Category", "n", "city") %>% 
  distinct() %>% 
  filter(Category == "Bird" )

new_bio_bird[new_bio_bird$Park.Name == "Death Valley", "State"] = "California"

# mammal
new_bio_mammal <- bio %>% 
  select("Park.Code", "Park.Name", "State","Latitude", "Longitude", "Category", "n", "city") %>% 
  distinct() %>% 
  filter(Category == "Mammal" )

new_bio_mammal[new_bio_mammal$Park.Name == "Death Valley", "State"] = "California"

```



#### 1. biodiversity & visitaion 2017
Tried to use the data generated above, but it's hard to draw two legend with 3 different data. Thus, I transformed and combined the three data in to one by rows. Multiply mammal and bird amount by 1000 to ensure color show the correct relationship between biodiversity and visitation, since biodiversity values are in 100 magnitude but visitation values are in 100k magnitude.
Here we use different shapes indicate bird, mammal and visitation, while the color refers to the number of each value. Red means more and blue means less.


##### final map just for visitition
```{r, warnings = FALSE}
us_map <- get_map("Unitd States", zoom = 4, maptype = "terrain")


# reshape the df for drawing plot
park_vis_17_3 <- park_vis_17 %>% filter(State %in% c("CA","CO","FL")) %>% 
  mutate( Category = "Park") %>% 
  rename(Park.Code = UnitCode, n = annual_vis) %>% na.omit() %>% 
  ungroup() %>%
  select(-c(Year, Acres, ParkName)) 

park_vis_17_3[park_vis_17_3$State == "CA","State"] = "California"
park_vis_17_3[park_vis_17_3$State == "CO","State"] = "Colorado"
park_vis_17_3[park_vis_17_3$State == "FL","State"] = "Florida"

bio_bird_all <- new_bio_bird %>% filter(State %in% c("California","Colorado","Florida")) %>% na.omit() %>% select(-c(city))
bio_mammal_all <- new_bio_mammal%>% filter(State %in% c("California","Colorado","Florida")) %>% na.omit()%>% select(-c(city))
all_data <- park_vis_17_3 %>% 
  rbind(bio_bird_all) %>% 
  rbind(bio_mammal_all)


label_data <- all_data %>% filter(Category == "Mammal")

park_data = all_data %>% filter(Category == "Park")

p_park <- ggmap(us_map) +
  scale_x_continuous( limits = c(-126 , -78 ), expand = c( 0 , 0 )   )+
  scale_y_continuous( limits = c( 22.5 , 43 ), expand = c( 0 , 0 ) )+
  geom_point(data = park_data, aes(x = Longitude, y = Latitude, colour = n), size = 4)+
  scale_colour_gradient(low = "blue", high = "red", name="visits",breaks = c(min(park_data$n),max(park_data$n)),labels = c("4.5k","370k"))+
  geom_label_repel(data = label_data, aes(x = Longitude, y = Latitude), 
                   label = label_data$Park.Name, size = 3,alpha = 1, 
                   label.size = NA, label.padding=0.07, force = 20 )+
  theme_void() + 
  ggtitle("Park Visitation for all 3 states") 

p_park

# ggsave("pak_vis.png", dpi = "retina")



```

##### final map just for biodiversiy
```{r, warnings = FALSE, message=FALSE}

all_data <- park_vis_17_3 %>% 
  rbind(bio_bird_all) %>% 
  rbind(bio_mammal_all)

all_data$Longitude[all_data$Category == "Mammal"] <- all_data$Longitude[all_data$Category == "Mammal"]+1


label_data <- all_data %>% filter(Category == "Mammal")

bio_data <- all_data %>% filter(Category != "Park")

p_park <- ggmap(us_map) +
  scale_x_continuous( limits = c(-126.5 , -77 ), expand = c( 0 , 0 )   )+
  scale_y_continuous( limits = c( 22.5 , 43 ), expand = c( 0 , 0 ) )+
  geom_point(data = bio_data, aes(x = Longitude, y = Latitude, colour = n, shape = Category), size = 4)+
  scale_colour_gradient(low = "blue", high = "red", name="number")+
  geom_label_repel(data = label_data, aes(x = Longitude, y = Latitude), 
                   label = label_data$Park.Name, size = 3,alpha = 1, 
                   label.size = NA, label.padding=0.07, force = 20 )+
  # geom_text_repel(data = label_data, aes(x = Longitude, y = Latitude, label = Park.Name), size = 3.5) +
  theme_void() + 
  scale_shape_manual(values=c(17,15)) +
  ggtitle("Biodiversity for all 3 states") 

p_park

# ggsave("bio_map.png", dpi = "retina")
```
